An open-source integrated framework for the automation of citation collection and screening in systematic reviews.

### 2022-02-19 Manuscript/Manuscript.tex
English proof reading and better number formatting

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{caption}
\usepackage{placeins}
\usepackage{amsmath}
\usepackage{lineno}
\linenumbers
\captionsetup[table]{labelformat=empty}
\captionsetup[figure]{labelformat=empty}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}
The exponential and prolific growth of published original scientific
contributions makes secondary literature abridgements increasingly
demanding. We introduce a new open-source framework for systematic
reviews that significantly reduces time and human resource allocation
for text identification, collection, and screening of scientific
literature.\\
The framework introduces three main tools: 1) an automatic citation
search engine and manager that allows collecting records from multiple
citation databases (Pubmed, WOS, SCOPUS, EMBASE, IEEE) with a unified
query syntax; 2) a citation screening tool based on Bayesian active
machine learning and natural language processing, which requires users
feedback only on uncertain classifications to iteratively increase
predictive accuracy; 3) a semi-automatic data-driven query generator to
create new search queries from existing reviewed citation data sets.\\
The framework was tested on an example topic to perform citation
collection and screening of related scientific production. To evaluate
the performance of the active machine learning classifier, we estimated
the median posterior Sensitivity and Efficiency {[}90\% Credible
Intervals{]} using Bayesian simulation to predict the distribution of
potentially relevant matches among the records not manually reviewed.\\
17755 unique records were collected through the framework citation
manager; 101 over 766 records were found to be relevant after manual
evaluation, while the rest were excluded by automatic classification;
the expected Efficiency was 95.7\% {[}95.3\%, 95.7\%{]} with a
Sensitivity of 100\% {[}93.5\%, 100\%{]}.\\
An additional search query generated from the labelled dataset,
retrieved 82579 additional records of which only 567 records required
human review and six additional positive matches were retrieved.
Including the additional records, the overall expected Sensitivity
decreased to 97.3\% {[}73.8\%, 100\%{]} while the Efficiency achieved
was 98.6\% {[}98.2\%, 98.7\%{]}.\\
For large studies, the developed framework significantly reduces
workload, human effort and error, simplifying citation collection and
screening with exceptional sensitivity, hence improving the
standardization and repeatability of systematic reviews.
\end{abstract}

@@

is especially true for biomedical research, a trend further increased by
the COVID-19 pandemic, thanks to faster article' processing time by
publishers and the diffusion of preprint databases' usage (Aviv-Reuven
and Rosenfeld 2021; Horbach 2020; Hoy 2020). Consequently, it gets
harder for researchers and practitioners to stay up-to-date on the

@@

relevance in this scenario, providing valuable summaries of the latest
research results, but is getting ever more demanding in terms of time
and human resources (Allen and Olkin 1999; Borah et al. 2017; A. M.
Cohen et al. 2010; Bastian, Glasziou, and Chalmers 2010).\\

@@

research needs to be collected from scientific databases through
appropriately built search queries (retrieval phase); secondly, the
acquired scientific citations need to be screened, selecting only those

@@

The construction of search queries is a complex task (Lefebvre et al.
2011; Hammerstrøm et al. 2010), requiring both domain and some knowledge
of the databases' query languages; the goal is to produce a set of
results containing all relevant articles (high sensitivity) while
keeping the total number low (high specificity), focusing on the first
aspect at the cost of the second (Hammerstrøm et al. 2010). If an
integrated search tool is not used, manual work is required to download,
store and organise the publication data; this approach is complicated by
limits in the number of records that can be downloaded at once and the
necessity to harmonise different formats and resolve record duplication
(Marshall and Wallace 2019).\\

@@

of a systematic review: even with appropriately built search queries,
the returned results easily range in the tens of thousands of which just
a small fraction is actually relevant (Lefebvre et al. 2011). It was
estimated that labelling 10,000 publications may take as much as 40
weeks of work and that the average clinical systematic review takes 63
weeks to be completed (Bannach-Brown et al. 2019; Borah et al. 2017;
Allen and Olkin 1999). A consequence is that often systematic reviews
are already outdated once they are published (E. M. Beller et al.
2013).\\

@@

Beller et al. 2018; Tsafnat et al. 2014). Through the application of
natural language processing (NLP), it is possible to transform free text
into quantitative features, with various levels of abstraction and

@@

2008); with machine learning, such text-derived data can be used to map
and reproduce human judgment, automating the citation screening

@@

The automation of systematic reviews has been ripe with improvements in

@@

to foresee that it is going to become the standard approach in the field
(E. Beller et al. 2018), with many solutions already being turned into
commercial and free-to-use tools (see Marshall and Wallace 2019, table

@@

In this manuscript we present an open source, production-ready framework
which further contributes to the state-of-the-art in systematic review
automation (SRA) and helpers (SRH) tools. We improve the ``retrieval
phase'' by providing a unified framework for the automatic collection
and management of scientific literature from multiple online sources.
For the citation screening (appraisal) phase, we built an active machine
learning-based Miwa et al. (2014) protocol, which exploits a Bayesian
framework to efficiently identify potentially relevant documents that
require human review while automatically screening-out the large
majority of clearly non-relevant ones; the algorithm then uses human
reviews to iteratively increase classification accuracy. Finally, we
included a tool to generate new search queries given an already labelled
citation data set, in order to identify relevant research possibly
missed by human-made queries.\\

@@

example topic of interest for our group: the evaluation of the

@@

We give an overview of the framework in the Methods section; in the
Result section, we show the framework outputs and performance once
applied to the example topic; in the Discussion, we lay out the
methodological justification behind the different components and
features of the framework.\\

@@

We built an R (R Core Team 2020) based framework to simplify two aspects
of systematic reviews: record acquisition and classification. The code
used to generate the results in the manuscript is available at X while
an update version of the framework is distributed as an R package at Y.
The framework includes several modules that communicate through
intermediate outputs stored in standard formats, which make it possible
for users to extend the framework or easily integrate it with other
tools in their pipeline. See Supplemental Material S1 for an in-depth
description of the framework and how to use it.\\
The tasks carried out by the framework are grouped into ``sessions'',
which comprise obtaining scientific citation data (records) using a
search query and then labelling them as relevant (``positive'' in the
rest of the text) or not (``negative'') for the topic of interest with
the help of a machine learning engine (Fig. 1). The initial search query
should be built using domain knowledge, trying to achieve a high
relevant/non-relevant record ratio.\\
The framework can then generate a new data-driven query from this
labelled set to perform a new session to find records possibly missed by
the first query.\\

@@

\hypertarget{records-acquisition-and-initial-labelling}{%
\subsection{Record's acquisition and initial
labelling}\label{records-acquisition-and-initial-labelling}}

We built a set of tools to let users automatically search and download
citation data from three major scientific databases (``sources''):
Pubmed (\url{https://pubmed.ncbi.nlm.nih.gov/}), Web Of Science (WOS,
\url{https://apps.webofknowledge.com/}) and the Institute of Electrical
and Electronics Engineers (IEEE,
\url{https://ieeexplore.ieee.org/Xplore/home.jsp}). The framework takes
care of authorization management for non-open databases like WOS and
IEEE. It is also possible to download and import records in the
framework manually; this is particularly useful to acquire records from
the SCOPUS (\url{https://www.scopus.com/}) and EMBASE databases
(\url{https://www.embase.com/}), for which a comprehensive API interface
was not easy to build. An extra manual search was necessary also for
Pubmed, since the API and the web interface have different rule
expansion algorithms and return slightly different results ({`NCBI
Insights : Updated Pubmed e-Utilities Coming in April 2022!'}, n.d.). A
short guide on how to set up the framework for each supported database
is available in Supplemental Material S3.\\
The acquired records are merged into a single database, resolving
duplicates and different formatting between sources. The records are
ordered according to the frequency of the positive query terms (e.g.,
not preceded by a \(NOT\) modifier) in the title and abstract (``simple
query ordering'').\\
The researcher is then asked to label a number of records to create the
``initial training set'' needed to start the automatic classification.
We suggest manually labelling the first 250 records (see
``hyperparameter optimization'' later). The simple query ordering
increases the positivity rate in the initial training set (Wallace,
Small, et al. 2010), which provide higher sensitivity during automatic
classification (Chawla, Japkowicz, and Kotcz 2004).

@@

The collected citation data have a number of fields characterising a
scientific publication. The framework models the relevance of a record
based on the following fields: title, abstract, authors, keywords, MESH
terms (Lipscomb 2000). A series of Natural Language Processing (NLP)
techniques (Baeza-Yates, Ribeiro-Neto, et al. 1999; Marshall and Wallace
2019; Ananiadou and McNaught 2006) are employed to transform the textual
information in these fields into features for machine learning through a
bag-of-words approach (Marshall and Wallace 2019). The processing of
free text fields (title, abstract) includes: tokenization (i.e.,
extracting the terms), common stopwords (i.e., sentence components
bringing no meaning) removal, part-of-speech filtering (only nouns,
adjectives, verbs and untagged terms are retained), and lemmatization of
the terms (i.e., reduction to their base grammar form). Text processing
for authors, keywords and MESH terms identifies logical units (e.g.,
author's full names, composite keywords) and extracts them.\\
Terms appearing in less than 5\% of the labelled documents are removed
from negative records. All terms in the positive set are kept to
increase sensitivity at the cost of specificity.\\
Some terms tend to co-appear in records (non-consecutive ngrams,
nc-ngrams), often carrying a particular meaning when copresent. To
detect nc-ngrams, we generated a word network representation (Francois
Rousseau 2015) posing edges between terms with a cosine similarity in
terms of record co-occurrence \textgreater{} 0.5. We extracted the
maximal cliques in the network (Eppstein, Löffler, and Strash 2010)
representing highly correlated groups of terms; These groups of terms
are added to the data set as individual features. Only nc-ngrams
comprising a maximum of ten terms are kept.\\
A second network is built using a co-occurrence threshold of 0.9. In
this case, the cliques represent terms that always appear together and
therefore can be considered redundant (i.e., they do not need to be
considered separately). These terms are merged to increase computation
efficiency and reduce overfitting.\\
The output is a Document-Term Matrix (DTM), with \(N_d\) rows
representing the records (\(D_i\)), \(N_t\) terms column for the
\(t_{field}\) terms (divided by record field) and \({0,1}\) values
whether \(t_{field} \in D_i\). We also enriched the DTM with features
referencing the number of terms in each field to help the model scale
term importance based on the field length.

@@

We used a Bayesian Additive Regression Trees (BART) machine learning
``classification model'' (Chipman et al. 2010) (in the implementation of
Kapelner and Bleich 2013) to predict the probability of a record of
being relevant, given the information coded into the enriched DTM and
the initial training set. We set up the BART model to use 2000 MCMC
iterations (after 250 burn-in iterations) and 50 trees; we used a \(k\)
value of 2 to regularize extreme prediction and let the model use
missing fields in the DTM as features (Kapelner and Bleich 2015).
Positive records are oversampled ten times to increase sensitivity
(Batista, Prati, and Monard 2004).\\
The output is a posterior predictive distribution (PPD) for each record
describing its probability of being relevant (i.e., a positive match).
An ensemble of ten models was fitted to improve prediction stability by
averaging the PPD between models (Zhou 2021; Dietterich 2000).\\

To assign the labels, we employed an ``active learning'' Miwa et al.
(2014) approach, where a human reviews a specific subset of predictions
made from the machine, which is then retrained on the manually reviewed
data set. This process proceeds iteratively, decreasing prediction
uncertainty.\\
Label assignment is done through the identification of an ``uncertainty
zone'' whose construction is possible thanks to the Bayesian nature of
BART, which provides full PPDs instead of point-wise predictions for
each record.\\
To describe the process formally, we define

@@

as respectively the lower and upper boundaries of the 98\% quantile
interval of \(\pi_i\) (98\% Predictive Interval, 98\% PrI).\\
Then we identify the ``uncertainty zone'' as

@@

that is, the record lower 98\% PrI boundary should be higher than every
value in the uncertainty zone. In other words, for a record to be
labelled positive, its PPD should be within the range of the mixture of
PPD of the previously labelled positive records and not cross the
distributions of the negative records.\\
Conversely, a record is labelled as negative if

@@

Manual review is then necessary for: 1) uncertain records, 2) positive
records (to avoid false positives), 3) records whose predicted label
differs from the existing manual one. The last case helps identify human
errors or inconsistent labelling criteria.

The automatic classification and the manual review steps alternate in a
loop (CR iterations) until no new positive matches are found in four
consecutive iterations.

@@

BART model it is possible to extract the proportion of times a term was
used in a posterior tree over the sum of total inclusions of all
variables (Kapelner and Bleich 2013), as a measure of feature importance
for the model. We extracted the terms, the part of the citation data
they were used in, the average ``inclusion rate'' among the ensemble
models (over 10,000 inclusions) and its ratio over the standard
deviation of this inclusion (inclusion stability, IS). For each term we
ran a Poisson regression to get its linear association with a positive
label and reported it as Relative Risk (RR) with the number of standard
errors as significance index (Statistic); the comparison between the
inclusion rate in the BART models and the linear association allows to
spot relevant non-linear effects (i.e, the feature is relevant only in
association with others). We reported in the Results only the first (in
order of inclusion rate) 15 terms with IS \textgreater{} 1.5 while the
first fifty terms regardless of inclusion stability are reported in
Supplemental Material S2.

@@

We created an algorithm that generates a new search query to acquire
further relevant publications missed during the first search, possibly
at a reasonable cost in specificity (i.e., a higher number of negative
results).\\
The algorithm encompasses a number of steps:

\begin{itemize}
\tightlist
\item
  We fit a partition tree (Therneau and Atkinson 2019) between the DTM
  and 800 samples from the PPD; if a term is present multiple times in
  the DTM (e.g., both title and abstract), they are counted just one,
  and field term count features are removed. This step generates a list
  of rules composed by \(AND\)/\(NOT\) ``conditions'' made of
  terms/authors/keywords/MESH tokens, which together identify a group of
  records.
\item
  For each rule, negative conditions (i.e., \(NOT\) statements) are
  added iteratively, starting from the most specific one, until no
  conditions are found that would not also remove positive records.
\item
  The extended set of rules is sorted by positive-negative record
  difference in descending order. The cumulative number of unique
  positive records is computed and used to group the rules. Rules inside
  each group are ordered by specificity.
\item
  The researcher is then asked to review the rule groups, selecting one
  or more rules (useful if they convey different meaning) from each, or
  edit them (in case too specific positive or negative conditions were
  included). It is possible to exclude a group of rules altogether,
  especially those with the worse sensitivity/specificity ratio.
\item
  The selected rules are joined together by \(OR\) statements, defining
  a subset of records with a sensibly higher proportion of positive
  records than the original one.
\item
  Redundant rules (i.e., rules whose positive records are already
  included in more specific ones) and conditions (i.e., conditions that
  once removed do not decrease the total number of positive or do not
  increase the negative records) are removed.
\item
  Finally, the rules are re-elaborated in a query usable on the major
  scientific databases.
\end{itemize}

Since the algorithm is data-driven, it creates queries that effectively
select positive records from the input data set but may be not specific
enough once applied to actual research databases. Therefore we appended
an extra subquery in \(AND\), which specifies the general topics of our
search and delimitates the search domain.\\
The new query was used to initiate a second search session.

@@

\begin{itemize}
\tightlist
\item
  Size of the initial training set: 50, 100, 250, 500 records;
\item
  Number of models in the ensemble: 1, 5, 10, 20, 40, 60 repetitions;
\item
  Oversampling rate of positive records: (1x (i.e., no oversampling),
  10x, 20x;
\item
  PrI quantiles for building the uncertainty zone: 80\%, 90\%, 98\%;
\item
  Source of randomness between models in the ensemble: MCMC sampling
  only (Robert, Casella, and Casella 2004), MCMC plus data bootstrapping
  (Breiman 1996) of the training set.
\end{itemize}

>>>>

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{caption}
\usepackage{placeins}
\usepackage{amsmath}
\usepackage{lineno}
\linenumbers
\captionsetup[table]{labelformat=empty}
\captionsetup[figure]{labelformat=empty}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

\begin{abstract}
The exponential and prolific growth of published original scientific
contributions makes secondary literature abridgements increasingly
demanding. We introduce a new open-source framework for systematic
reviews that significantly reduces time and human resource allocation
for text identification, collection, and screening of scientific
literature.\\
The framework provides three main tools: 1) An automatic citation search
engine and manager that allows collecting records from multiple citation
databases (Pubmed, WOS, SCOPUS, EMBASE, IEEE) with a unified query
syntax 2) A citation screening tool based on Bayesian active machine
learning and natural language processing that requires users feedback
only on uncertain classifications to iteratively increase predictive
accuracy and 3) A semi-automatic, data-driven query generator to create
new search queries from existing reviewed citation data sets.\\
The framework was tested on an example topic to collect and screen
citations of related scientific production. To evaluate the performance
of the active machine learning classifier, we estimated the median
posterior sensitivity and efficiency {[}90\% Credible Intervals{]} using
Bayesian simulation to predict the distribution of potentially relevant
matches among the records not manually reviewed.\\
A total of 17,755 unique records were collected through the framework
citation manager; 101 of 766 records were found to be relevant after
manual evaluation, while the rest were excluded by automatic
classification; the theorethical efficiency was 95.6\% {[}95.3\%,
95.7\%{]} with a sensitivity of 100\% {[}93.5\%, 100\%{]}.\\
A new search query was generated from the labelled dataset, and 82,579
additional records were collected; only 567 records required human
review after automatic screening, and six additional positive matches
were found. Including the additional records, the overall expected
sensitivity decreased to 97.3\% {[}73.8\%, 100\%{]} while the efficiency
increased to 98.6\% {[}98.2\%, 98.7\%{]}.\\
For large studies, the framework can significantly reduce the human
resources required to conduct systematic reviews by simplifying citation
collection and screening while demonstrating exceptional sensitivity.
Such tool can improve the standardization and repeatability of
systematic reviews.
\end{abstract}

@@

is especially true for biomedical research, a trend further accelerated
by the COVID-19 pandemic, thanks to faster article' processing time by
publishers and the greater use of preprint databases (Aviv-Reuven and
Rosenfeld 2021; Horbach 2020; Hoy 2020). Consequently, it has become
harder for researchers and practitioners to stay up to date on the

@@

importance in this scenario in that it provides valuable summaries of
the latest research results; however, it is becoming ever more
challenging in terms of time and human resources required (Allen and
Olkin 1999; Borah et al. 2017; A. M. Cohen et al. 2010; Bastian,
Glasziou, and Chalmers 2010).\\

@@

research must be collected from scientific databases using appropriately
built search queries (retrieval phase); secondly, the scientific
citations collected must be screened, selecting only those that are

@@

Search queries construction is a complex task (Lefebvre et al. 2011;
Hammerstrøm et al. 2010), requiring both expertise in the scientific
field of interest and some knowledge of the database query languages.
The goal is to obtain a set of results that contains all relevant
articles (high sensitivity) while keeping the total number of records
low (high specificity), possibly focusing on the first at the expense of
the second (Hammerstrøm et al. 2010).\\
If an integrated search tool is not used, manual work is required to
download, store and organise the publication data; this approach is
complicated by limits to the number of records that can be downloaded at
any one time and the need to harmonise different formats and resolve
record duplication (Marshall and Wallace 2019).\\

@@

in a systematic review: even with appropriately built search queries,
the results may easily range in the tens of thousands, of which just a
small fraction are actually relevant (Lefebvre et al. 2011). It has been
estimated that labelling 10,000 publications can take up to 40 weeks of
work and that the average clinical systematic review takes 63 weeks to
complete (Bannach-Brown et al. 2019; Borah et al. 2017; Allen and Olkin
1999). A consequence of this is systematic reviews are often already
out-of-date by the time they are published (E. M. Beller et al. 2013).\\

@@

Beller et al. 2018; Tsafnat et al. 2014). By applying natural language
processing (NLP), it is possible to transform free text into
quantitative features, with various levels of abstraction and

@@

2008); using machine learning, such text-derived data can be used to map
and reproduce human judgment, automating the screening of citations

@@

Automation of systematic reviews has made significant improvements in

@@

foreseeable that it will become the standard approach in the field (E.
Beller et al. 2018), with many solutions already being implemented into
commercial or free-to-use tools (see Marshall and Wallace 2019, table

@@

In this manuscript, we present an open source, production-ready
framework that further contributes to the state-of-the-art in systematic
review automation (SRA) and helpers (SRH) tools. We improve the
``retrieval phase'' by providing a unified framework for the automated
collection and management of scientific literature from multiple online
sources. For the citation screening (appraisal) phase, we built an
active machine learning-based protocol (Settles 2009; Miwa et al. 2014),
which utilises a Bayesian framework to efficiently identify potentially
relevant documents that require human review, while automatically
screening-out the vast majority of clearly non-relevant ones; the
algorithm then requires human review to iteratively increase
classification accuracy. Finally, we included a tool to generate new
search queries based on an already labelled citation data set, to
identify relevant research that may possibly have been missed by
manually-made queries.\\

@@

example topic of interest to our group: the evaluation of the

@@

In the Methods, we give an overview of the framework, in the Result, we
show the outputs and performance of the framework applied to the example
topic, and in the Discussion, we explain the methodological rationale
for the different components and features of the framework.\\

@@

We built an R (R Core Team 2020) based framework to simplify two aspects
of systematic literature review: record acquisition and classification.
The code used to generate the results in the manuscript is available at
\url{https://github.com/AD-Papers-Material/BART_SystReviewClassifier},
while an updated and ready to use version of the framework is
distributed as an R package at
\url{https://github.com/bakaburg1/BaySREn}. The framework includes
several modules that communicate through intermediate outputs stored in
standard formats, which make it possible for users to extend the
framework or easily integrate it with other tools in their pipeline. See
Supplemental Material S1 for an in-depth description of the framework
and how to use it.\\
The tasks carried out by the framework are grouped into ``sessions'',
which comprise obtaining scientific citation data (records) using a
search query and then labelling them as relevant (``positive'' in the
rest of the text) or not (``negative'') for the topic of interest with
the help of a machine learning engine (Fig. 1). The initial search query
should be built using domain knowledge, trying to achieve a high
relevant/non-relevant record ratio.\\
The framework can then generate a new data-driven query from this
labelled set to perform a new session to find records possibly missed by
the first query.\\

@@

\hypertarget{record-acquisition-and-initial-labelling}{%
\subsection{Record acquisition and initial
labelling}\label{record-acquisition-and-initial-labelling}}

We built a set of tools to allow users to automatically search and
download citation data from three major scientific databases
(``sources''): Pubmed (\url{https://pubmed.ncbi.nlm.nih.gov/}), Web Of
Science (WOS, \url{https://apps.webofknowledge.com/}) and the Institute
of Electrical and Electronics Engineers (IEEE,
\url{https://ieeexplore.ieee.org/Xplore/home.jsp}). The framework takes
care of authorization management for non-open databases like WOS and
IEEE. It is also possible to download and import records in the
framework manually; this is particularly useful to acquire records from
the SCOPUS (\url{https://www.scopus.com/}) and EMBASE databases
(\url{https://www.embase.com/}), for which a comprehensive API interface
was not easy to build. An extra manual search was also necessary for
Pubmed, since the API and the web interface have different rule
expansion algorithms and return slightly different results ({`NCBI
Insights : Updated Pubmed e-Utilities Coming in April 2022!'}, n.d.). A
short guide on how to set up the framework for each database supported
is available in Supplemental Material S3.\\
The collected records are merged into a single database, resolving
duplicates and different formatting between sources. The records are
ordered according to the frequency of the positive query terms (e.g.,
not preceded by a \emph{NOT} modifier) in the title and abstract
(``simple query ordering'').\\
The researcher is then asked to label a number of records to create the
``initial training set'' needed to start the automatic classification.
We recommend manually labelling the first 250 records (see
``hyperparameter optimization'' later). Simple query ordering increases
the positivity rate in the initial training set (Wallace, Small, et al.
2010), leading to higher sensitivity during automatic classification
(Chawla, Japkowicz, and Kotcz 2004).

@@

The collected citation data contain a number of fields characterising a
scientific publication. The framework models the relevance of a record
based on the following fields: title, abstract, authors, keywords, MESH
terms (Lipscomb 2000). A range of Natural Language Processing (NLP)
techniques (Baeza-Yates, Ribeiro-Neto, et al. 1999; Marshall and Wallace
2019; Ananiadou and McNaught 2006) are employed to convert the textual
information in these fields into features for machine learning through a
bag-of-words approach (Marshall and Wallace 2019). Processing of free
text fields (title, abstract) includes: tokenization (i.e., extracting
the terms), removal of common stopwords (i.e., sentence components
having no semantic value), part-of-speech filtering (only nouns,
adjectives, verbs and untagged terms are retained), and lemmatisation of
terms (i.e., reduction to their base grammatical form). Text processing
for authors, keywords and MESH terms identifies logical units (e.g.,
authors' full names, composite keywords) and extracts them.\\
Terms appearing in less than 5\% of the labelled documents are removed
from negative records. All terms in the positive set are kept to
increase sensitivity at the cost of specificity.\\
Some terms tend to co-appear in records (non-consecutive ngrams,
nc-ngrams), often carrying a particular meaning when they do co-occur.
To detect nc-ngrams, we generated a word network representation
(Francois Rousseau 2015) with edges occurring between terms with a
cosine similarity in terms of docuement co-occurrence \textgreater{}
0.5. We extracted the maximal cliques in the network (Eppstein, Löffler,
and Strash 2010) representing highly correlated groups of terms; these
groups are added to the dataset as individual features. Only nc-ngrams
comprising a maximum of ten terms are kept.\\
A second network is built using a co-occurrence threshold of 0.9. In
this case, the cliques represent terms that always appear together and
can therefore be considered redundant (i.e., they do not need to be
considered separately). These terms are merged to increase computational
efficiency and reduce overfitting.\\
The output is a Document-Term Matrix (DTM), with \(N_d\) rows
representing the records (\(D_i\)), \(N_t\) terms column for the
\(t_{field}\) terms (divided by record field) and \({0,1}\) values
whether \(t_{field} \in D_i\). We also enriched the DTM with features
referencing the number of terms in each field to help the model scale
term importance based on the field length.

@@

We used a Bayesian Additive Regression Trees (BART) machine learning
``classification model'' (Chipman et al. 2010) (in the implementation of
Kapelner and Bleich 2013) to predict the probability of a record being
relevant, given the information coded into the enriched DTM and the
initial training set. We set up the BART model to use 2,000 MCMC
iterations (after 250 burn-in iterations) and 50 trees; we used a \(k\)
value of 2 to regularize extreme prediction and let the model use
missing fields in the DTM as features (Kapelner and Bleich 2015).
Positive records are oversampled ten times to increase sensitivity
(Batista, Prati, and Monard 2004).\\
The output is a posterior predictive distribution (PPD) for each record
describing the probability of it being relevant (i.e., a positive
match). An ensemble of ten models was fitted to improve prediction
stability by averaging the PPD between models (Zhou 2021; Dietterich
2000).\\

To assign the labels, we employed an ``active learning'' approach
(Settles 2009; Miwa et al. 2014), where a human reviews a specific
subset of predictions made by the machine, which is then retrained on
the manually reviewed dataset. This process is carried out iteratively
to reduce prediction uncertainty.\\
Label assignment is done through identification of an ``uncertainty
zone'' the construction of which is possible thanks to the Bayesian
nature of BART, which provides full PPDs instead of point-wise
predictions for each record.\\
To describe the process formally, we define

@@

respectively as the lower and upper boundaries of the 98\% quantile
interval of \(\pi_i\) (98\% Predictive Interval, 98\% PrI).\\
Then we identify the ``uncertainty zone'' as

@@

that is, the record lower 98\% PrI boundary should be higher than every
value in the uncertainty zone. In other words, for a record to be
labelled positive, its PPD should be within the range of the mixture of
PPD of the records previously labelled positive, and should not cross
the distributions of the negative records.\\
Conversely, a record is labelled as negative if

@@

Manual review is then necessary for: 1) uncertain records, 2) positive
records (to avoid false positives), and 3) records whose predicted label
differs from the existing manual one. The last case helps identify human
errors or inconsistent labelling criteria.

The automatic classification and manual review steps alternate in a loop
(CR iterations) until no new positive matches are found in four
consecutive iterations.

@@

With the BART model, it is possible to extract the proportion of times a
term is used in a posterior tree over the sum of total inclusions of all
variables (Kapelner and Bleich 2013), as a measure of importance of the
features in the model. We extracted the terms, the portion of the
citation data in which they were used, the average ``inclusion rate''
among the ensemble models (over 10,000 inclusions) and its ratio over
the standard deviation of this inclusion (inclusion stability, IS). For
each term we ran a Poisson regression to get the linear association with
a positive label and reported it as Relative Risk (RR) with the number
of standard errors as significance index (Statistic); the comparison
between the inclusion rate in the BART models and the linear association
allows to spot relevant non-linear effects (i.e, the feature is relevant
only in association with others). In the Results, we only listed the
first 15 terms with IS \textgreater{} 1.5, (in order of inclusion rate),
while the first fifty terms regardless of inclusion stability are listed
in Supplemental Material S2.

@@

We developed an algorithm that generates a new search query to find
further relevant publications missed in the initial search, possibly at
a reasonable cost to specificity (i.e., a higher number of negative
results).\\
The algorithm encompasses a number of steps:

\begin{itemize}
\tightlist
\item
  A partition tree (Therneau and Atkinson 2019) is built between the DTM
  and 800 samples from the PPD; if a term is present multiple times in
  the DTM (e.g., both in the title and abstract), it is counted just
  once, and field term count features are removed. This step generates a
  list of rules composed by \emph{AND}/\emph{NOT} ``conditions'' made of
  terms/authors/keywords/MESH tokens, which together identify a group of
  records.
\item
  For each rule, negative conditions (i.e., \emph{NOT} statements) are
  added iteratively, starting from the most specific one, until no
  conditions are found that would not also remove positive records.
\item
  The extended set of rules is sorted by positive-negative record
  difference in descending order. The cumulative number of unique
  positive records is computed and used to group the rules. Rules inside
  each group are ordered by specificity.
\item
  The researcher is then asked to review the rule groups and select one
  or more rules from each group (useful if they convey different
  meaning), or edit them if overly too specific positive or negative
  conditions were included. It is possible to exclude a group of rules
  altogether, especially those with the poorest sensitivity/specificity
  ratio.
\item
  The selected rules are joined together by \emph{OR} statements,
  defining a subset of records with a sensibly higher proportion of
  positive records than the original set
\item
  Redundant (i.e., rules whose positive records are already included in
  more specific ones) and non-relevant rules (i.e., conditions that when
  removed do not impact sensitivity and specificity) are removed.
\item
  Finally, the rules are re-elaborated in a query that can be used in
  the major scientific databases.
\end{itemize}

Because the algorithm is data-driven, it creates queries that
effectively select positive records from the input dataset, but may be
not specific enough when applied to actual research databases. Therefore
we added an extra subquery in \emph{AND}, that specifies the general
topics of our search and narrows the search domain.\\
The new query was used to initiate a second search session.

@@

\begin{itemize}
\tightlist
\item
  Size of the initial training set: 50, 100, 250, 500 records;
\item
  Number of models in the ensemble: 1, 5, 10, 20, 40, 60 repetitions;
\item
  Oversampling rate of positive records: (1x (i.e., no oversampling),
  10x, 20x;
\item
  PrI quantiles for building the uncertainty zone: 80\%, 90\%, 98\%;
\item
  Source of randomness between models in the ensemble: MCMC sampling
  only (Robert and Casella 2004), MCMC plus data bootstrapping (Breiman
  1996) of the training set.
\end{itemize}

To evaluate the hyperparameter effect of performance, we set up a ``grid
search'' (Claesen and De Moor 2015; L. Yang and Shami 2020) on a
prelabelled ``validation set'' derived from the first 1,200 records of
the first session dataset. Each hyperparameter combination was tested
until four CR iterations were completed with no positive records or the
whole dataset was labelled.\\
For each combination, a performance score was computed as the product of
``Efficiency'' (1 minus the ratio of records that required reviewing
over the total number of records) and ``Sensitivity'' (number of
positive records found over the total number of positive records). We
then used a partition tree (Therneau and Atkinson 2019) to identify
homogeneous ``performance clusters'' of scores given hyperparameter
values. For the rest of the study, we used the best hyperparameter set
in terms of Sensitivity followed by Efficiency from the cluster with the
highest average score.\\
